Reasoning About a Machine with Local Capabilities

### 2019-08-12 toplas/paper.tex
[TOPLAS] Add reviews with todo and first round of fixes

\emph{Capability machines} are a type of processors that remediate these limitations with a better security model at the hardware level.
They are based on old ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that have recently received renewed interest.
in particular, the CHERI project has proposed new ideas and ways of tackling practical challenges like backwards compatibility and realistic OS support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
Capability machines tag every word (in the register file and in memory) to enforce a strict separation between numbers and capabilities (a kind of pointers that carry authority).
Memory capabilities carry the authority to read and/or write to a range of memory locations.
There is also a form of \emph{object capabilities}, which represent the authority to invoke a piece of code without exposing the code's encapsulated private state (e.g., the M-Machine's enter capabilities (described in Section~\ref{sec:capab-mach-with}) or CHERI's sealed code/data pairs).

@@

\subsubsection{Future Worlds}
\label{subsec:future-worlds} 
The future world relations model how memory may evolve over time.
We have a \emph{public} and a \emph{private} future world relation that respectively what changes any capability can rely on and what changes only temporary regions can rely on.
As temporary capabilities fall within the category of all capabilities, the public future relation is subsumed in the private future relation.

@@

%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation defines semantically when values, program counters, and configurations are capability safe.
The logical relation is defined found in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds}, and we provide some explanations in the following paragraphs.
For space reasons, we omit some definitions and explain them only verbally, but precise definitions can be found in the appendix.
The logical relation is recursively defined, so we encourage first time readers to just read the section in its entirety and do a second read afterwards.

@@

Capabilities with read permission cannot directly break memory protocols because they cannot make changes to the memory.
However, a read capability could be used to read a capability that can break memory protocols.
For this reason, a capability with read permissions is only safe when it can be used only to read safe capabilities.
This is captured by the $\readCond{}$-condition. 
More precisely, the condition requires the memory addressed by the capability to actually be governed by some region $W(r)$.
This region should at least require all the memories accepted by the state interpretation function to contain safe words but stricter requirements may be imposed on the memories.
This is expressed in terms of an upper bound on the possibly permitted memories defined in terms of a so called ``standard region''.
In this case, it is a \tpwl{} $\iota^\pwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}). 
The \tpwl{} only accept memories with address range $[\start,\addrend]$.
More importantly, the \tpwl{}'s state interpretation function only accepts memory segments that only contain safe words.
The relation between the actual region in the world and the upper bound is $W(r) \nsubsim[n] \iota^\pwl_{[\start,\addrend]}$ (defined in Appendix~\ref{app:logical-relation}).
Essentially, the relation requires the two regions state transition systems to be the same. Further in the current state of region $W(r)$ and in any world, the state interpretation function should only allow memories that are also allowed in the standard region.

The locality of capabilities play a role in whether or not they are safe.
Generally speaking, global capabilities should only depend on permanent regions.
This is expressed in $\readCond{}$ with $\var{localityReg}(\gl,W)$ which projects out all the regions the capability can depend on based on its locality.
Specifically, if the capability is local, then all active regions (non-revoked) are project.
On the other hand, if the capability is global, then only the permanent regions are projected.

@@

\begin{toplas}
  In \sectionname~\ref{sec:stack-and-return-pointer}, we require capabilities with write-local permission to be local.
  This indicates that our logical relation should imply the same, namely that capabilities with write-local permission are local.
  \begin{lemma}[Stack capabilities are local]~
    \label{lem:wl-local}
    \begin{itemize}
    \item If $\memSat{\ms}{W}$ and $\npair{\stdcap} \in \stdvr(W)$ and $\start
      \leq \addrend$ and $\perm \in \{\rwlx,\rwl\}$, then $\gl = \local$.
    \end{itemize}
  \end{lemma}
  In our definition of worlds, nothing prevents a world from having regions that are overlapping.
  This might seem like an issue the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
  In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness condition on the world as it prevents the regions from overlapping.
  The memory satisfaction assumption in Lemma~\ref{lem:wl-local} is there to ensure a well-formed world.
\end{toplas}

@@

  In addition to these syntactic requirements, we also specify standard behavioural expectations of malloc.
  Intuitively, we require that when malloc is invoked with a length argument, then it will return a capability for a fresh piece of memory of that size.
  It should be fresh in the sense that malloc has not already allocated any of that memory and will not do so in the future.
  Also, when invoked with a nonsensical length argument such as a negative integer or a capability, malloc should simply fail.
  However, formulating these requirements is harder than one might expect.
  The problem is that a realistic implementation of malloc needs to rely on internal state that changes after every invocation and relies on invariants on that state.
  We can only expect that malloc behaves according to its specification if its internal state satisfies the current state of the invariants in an executing system.
  We express this in terms of the semantic model from \sectionname~\ref{sec:logical-relation}.

@@

The macro \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} captures the
parts of the calling convention related to actually transferring control to
adversarial code. Specifically, it pushes the contents of the private registers,
$\overline{r_{\var{priv}}}$, to the stack and pushes the ``restoration'' code to
the stack. The restoration code is executed as the first thing upon return; it
restores the stack pointer and the old program counter. After the restoration
code is pushed to the stack, \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} adjusts the pc
to point to the first instruction after the jump and pushes it to the stack. The
\texttt{scall}macro constructs a protected return pointer from a stack pointer by
adjusting it to point to the first instruction of the return code and
encapsulating it by restricting its permission to $\entry$. Next, the \texttt{scall}
macro reduces the range of authority of the stack pointer to the unused part of
the stack and clears it(as discussed in
Section~\ref{sec:stack-and-return-pointer}). Finally, the \texttt{scall} macro clears
non-argument registers and jumps to $r$. Upon return after the restoration code
on the stack has been executed, the \texttt{scall} macro pops the restoration code from
the stack and restores the private state by popping it from the stack to the
appropriate registers. \figurename~\ref{fig:scall-impl} displays the
implementation of \texttt{scall} and the restoration code used by
\texttt{scall}. The implementation of \texttt{scall} uses some of the macros we
present next.

@@

\begin{toplas}
\section{Reasoning about programs on a capability machine}
\label{sec:reasoning}
% \itoplassug{
%   \begin{itemize}
%   \item A section with the lemmas we use for reasoning about programs
%     \begin{itemize}
%     \item lemma 58, scall works. Include description of how it handles the
%       common things in call and leaves the argumentation for callee code and
%       continuation.
%     \item lemma 59, malloc works. If this is included, we should probably also
%       include the malloc specification.
%     \item lemma 60, create closure works.
%     \item Include a description about how the FTLR is used to reason about
%       unknown code.
%     \item Mention that this is a step on the way towards having a program logic,
%       and then in the discussion discuss why a program logic for this language
%       is not what we need.
%     \end{itemize}
%   \end{itemize}
% }
There are many details to get right when programming in assembly.
These details carry over to proofs about assembly programs, so many of the proofs in \sectionname~\ref{sec:examples} about example programs are a bit cumbersome.
It is especially annoying when the same line of reasoning is applied in multiple places.
To mitigate this, we have define a number of lemmas that capture common reasoning patterns in these proofs.
In this section, we present the most central lemmas used to reason about assembly programs.

@@

% Many details to get right when reasoning about low-level programs.
Another common pattern in proofs about programs on our capability machine concerns the use of \texttt{scall}.
The following lemma captures the commonalities of reasoning about programs using \texttt{scall}.
For instance, setting up the local stack frame and constructing stack pointers and protected return pointers for the callee always amount to the same line of reasoning which is captured by the lemma.
From another point of view, it can be seen as a specification for \texttt{scall}.
\begin{lemma}[$\mathtt{scall}$ works]
  \label{lem:scall-works}
  For all $n \in \nats$, $\ms, \ms_\stk, \ms_\var{unused}, \ms_f \in \MemSegments$, $W
  \in \Worlds$, $\reg \in \Regs$, $r,\overline{r_{\mathit{arg}}},
  \overline{r_{\mathit{priv}}} \in \RegName$, and $c_\var{next} \in \Caps$,
  if
  \begin{enumerate}
  \item \label{item:scall:memsat} $\memSat[n]{\ms}{\revokeTemp{W}}$ 
  \item $\dom(\ms_f) \cap (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
  \item \label{item:scall:stack} $(\reg,\ms) \text{ is looking at }
    \mathtt{scall}\;r(\overline{r_{\mathit{arg}}},
    \overline{r_{\mathit{priv}}}) \text{ followed by }
    c_{\mathit{next}}$
  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
  \item \label{item:scall:hyp-callee} \emph{Hyp-Callee} For all  $\ms_\var{rec},
    \ms_\var{unused}', \ms'' \in \MemSegments$, $W' \in \Worlds$, $c_\var{ret},
    c_\stk \in \Caps$, and $\reg' \in \Regs$, if
    \begin{itemize}
    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{rec}}
      \uplus \ms_{\mathit{unused}}')$,
    \item $W' =
      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{rec}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$,
    \item $\memSat[n-1]{\ms''}{W'}$
    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
      \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk, r \mapsto \reg(r)]$ 
    \item $\npair[n-1]{c_{\mathit{ret}}} \in \stdvr(W')$
    \item $\npair[n-1]{c_\stk} \in \stdvr(W')$
    \end{itemize}

@@

In the common case, \texttt{scall} is used to reason about untrusted code that we only have general assumptions about.
As explained previously, we use the FTLR (Theorem~\ref{thm:ftlr}) to argue about unknown code; this is no exception.
That is, when we argue that we assumption~\ref{item:scall:hyp-callee} in the \texttt{scall} lemma satisfied, we use the FTLR along with the assumptions we have (e.g.\ linking table, malloc capability etc.).
In Hyp-Callee, we assume the memory is safe, so it suffices to show that the register-file contains safe values, which amounts to showing that the arguments in the call are safe.

@@

% We stress that these lemmas
% are conveniences and are not strictly necessary to reason about programs for the
% capability machine.
In principle, one could reason solely in the semantic model
% without the use of these lemmas, but that would lead to a lot of repetition.
\end{toplas}

@@

The assertion on line 8 of Figure~\ref{fig:awkward} should never fail.
From the code, this seems natural since \texttt{adv} does not have access to $x$.
Therefore after the second invocation on line 7, $x$ should still be in the state it was before that invocation.
However, \texttt{adv} could have access to $\textit{cl}_x$ and could invoke $\textit{cl}_x$ again within the second invocation of \texttt{adv}.
Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.
If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up on line 8 of Figure~\ref{fig:awkward} with $x$ set to $1$, causing the assertion to fail.
Without going into details, such an attack is perfectly possible if \texttt{adv} has access to a call/cc primitive (or equivalent).
In other words, if we are able to prove that the assertion will never fail, this attack, and other similar attacks, on well-bracketed control flow are adequately prevented.

@@

  The next part in the execution is an \texttt{scall}, so, according to \sectionname~\ref{sec:reasoning}, we should apply the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
  For the sake of presentation, we here sketch some of the things the \texttt{scall} lemma actually takes care of.
  Based on the stack pointer's permission, we know by Lemma~\ref{lem:wl-local} that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{cf.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
  This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked\footnote{As a side note, it is also important for the reasoning about $x$ that this is a private future world.}.
  In $W_2$, two new regions govern the stack.
  One of the two new regions governs the caller stack frame, i.e.\ the part of the stack that contains the contents of private registers and stack recovery code.
  The region that takes care of this is a temporary static region which ensures that our local stack frame remains unchanged during the execution of the callback.
  The other new region takes care of the unused part of the stack which we are going to let the callback use for its execution.
  This is taken care of by a standard region $\iota^{pwl}$-region which allows the callback to store any safe value on the stack.
  The callback is global, so it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we assume that the arguments for the invocation of $c_{f4}$ are safe).
  Notice that had the callback been local, this would not be the case, but it would also not be safe to invoke as it might be a stack pointer as discussed in \sectionname~\ref{sec:stack-and-return-pointer}.
  Note also that arguing that the memory satisfies $W_2$ when we invoke the callback only works because we cleared the stack entirely (including the unused part) before the invocation.
  Otherwise, it might contain local values that are only known to be safe in $W_1$, but for which we do not know that they will remain safe in the private future world $W_2$.
  This part corresponds to arguing about Hyp-Callee in the \texttt{scall} lemma.

@@

\section{Discussion}
\label{sec:discussion}
\itoplassug{Reviewer C, esop, would have liked a discussion of (a) what could be done on a machine with a smaller set of capabilities, e.g., no local capability (we could probably not have done anything as the stack pointer could be stored on the heap and reused at unintended times.) and (b) what could be done with a larger/stronger set of available capabilities (we could include a discussion of linear capabilities).}
\itoplassug{Reviewer C, esop, asks whether our attacker model is reasonable (see tex comment). Maybe we should include a short discussion of the attacker model.}
% p.18 "they are linked with an adversary, adv, that is allowed to allocate memory but has no other capabilities".  Is this a reasonable model of the adversary?  (Genuinely naive question.)
\itoplassug{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
% * Also, the local capabilities seem related to Boyland's "Borrowing"
%   e.g. that is now supported in Rust amongst other research languages
%   (Pony etc) also mentioned in the ECOOP paper.
% * the combination of E and L capabilities seems very interesting
%   indeed.  I wonder if e.g. they could be used to support strong
%   models of encapsulation e.g. like ownership types. (see eg LNCS
%   7850)
\noindent\textbf{Calling convention}\\
\emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow correctness, we do not prove a general theorem that shows this because it is not clear what such a theorem should look like.
Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, do not take into account temporal properties \itoplas{because they only take into account what capabilities can be available and not when they are used.
  Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than such formulations, although this is not made very explicit.
In later work, \citet{skorstengaard_stktokens_2019} provide a more explicit and useful definition of control-flow correctness and local state encapsulation.
The idea of their definition is to define a variant of a simple capability machine with a built-in stack as well as call and return instructions.
The call and return instructions use the built-in stack which gives well-bracketed control and local state encapsulation by construction.
They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to one without which means that the programs running on the capability machine without a built-in stack behaves as though there was a built-in stack.
Their full abstraction proof uses a logical relation, but it differs in a number of ways from the one we present.
Their logical relation is a binary cross language logical relation tailored specifically to prove the full abstraction result.
This means that it could not be used to prove program correctness results similar to ours.
Their calling convention is based on something called linear capabilities rather than local capabilities.
Linear capabilities offer a different kind of limited revocation.
\itoplassug{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}

>>>>

\emph{Capability machines} are a type of processors that remediate these limitations with a better security model at the hardware level.
They are based on old ideas~\citep{Carter:1994:HSF:195473.195579,Dennis:1966:PSM:365230.365252,shapiro_eros:_1999} that have recently received renewed interest.
In particular, the CHERI project has proposed new ideas and ways of tackling practical challenges like backwards compatibility and realistic OS support~\citep{Watson2015Cheri,Woodruff:2014:CCM:2665671.2665740}.
Capability machines tag every word (in the register file and in memory) to enforce a strict separation between numbers and capabilities (a kind of pointers that carry authority).
Memory capabilities carry the authority to read and/or write to a range of memory locations.
There is also a form of \emph{object capabilities}, which represent the authority to invoke a piece of code without exposing the code's encapsulated private state (e.g., the M-Machine's enter capabilities (described in Section~\ref{sec:capab-mach-with}) or CHERI's sealed code/data pairs).

@@

\subsubsection{Future Worlds}
\label{subsec:future-worlds} 
The future world relations model how memory may evolve over time.
We have a \emph{public} and a \emph{private} future world relation that, respectively, model the memory changes any capability can rely on and the memory changes only local capabilities can rely on.
As local capabilities fall within the category of all capabilities, the public future relation is subsumed in the private future relation.

@@

%subsection introduction
% \dominique{I would try to motivate the LR defs by saying that the LR is
%   constructed to express the FTLR and the FTLR says that if a piece of code only
%   has access to values respecting the invariants in a world, then it will
%   respect them itself too.} 
The logical relation defines semantically when values, program counters, and configurations are capability safe.
The logical relation is defined in \figurename{}s~\ref{fig:logrel} and~\ref{fig:perm-conds}, and we provide some explanations in the following paragraphs.
For space reasons, we omit some definitions and explain them only verbally, but precise definitions can be found in the appendix.
The logical relation is recursively defined, so we encourage first time readers to just read the section in its entirety and do a second read afterwards.

@@

Capabilities with read permission cannot directly break memory protocols because they cannot make changes to the memory.
However, a read capability could be used to read a capability that can break memory protocols.
For this reason, a capability with read permissions is only safe when it can be used only to read safe capabilities.
This is captured by the $\readCond{}$-condition. 
More precisely, the condition requires the memory addressed by the capability to actually be governed by some region $W(r)$.
This region should at least require all the memories accepted by the state interpretation function to contain safe words but stricter requirements may be imposed on the memories.
This is expressed in terms of an upper bound on the possibly permitted memories defined in terms of a so called ``standard region''.
In this case, it is a \tpwl{} $\iota^\pwl_{[\start,\addrend]}$ (defined in \figurename~\ref{fig:std-reg}). 
The \tpwl{} only accepts memories with address range $[\start,\addrend]$.
More importantly, the \tpwl{}'s state interpretation function only accepts memory segments that only contain safe words.
The relation between the actual region in the world and the upper bound is $W(r) \nsubsim[n] \iota^\pwl_{[\start,\addrend]}$ (defined in Appendix~\ref{app:logical-relation}).
Essentially, the relation requires the two regions state transition systems to be the same. Further in the current state of region $W(r)$ and in any world, the state interpretation function should only allow memories that are also allowed in the standard region.

The locality of capabilities play a role in whether or not they are safe.
Generally speaking, global capabilities should only depend on permanent regions.
This is expressed in $\readCond{}$ with $\var{localityReg}(\gl,W)$ which projects out all the regions the capability can depend on based on its locality.
Specifically, if the capability is local, then all active regions (non-revoked) are projected.
On the other hand, if the capability is global, then only the permanent regions are projected.

@@

\begin{toplas}
  In \sectionname~\ref{sec:stack-and-return-pointer}, we require capabilities with write-local permission to be local.
  This indicates that our logical relation should imply the same, namely that capabilities with write-local permission are local.
  \begin{lemma}[Stack capabilities are local]~
    \label{lem:wl-local}
    \begin{itemize}
    \item If $\memSat{\ms}{W}$ and $\npair{\stdcap} \in \stdvr(W)$ and $\start
      \leq \addrend$ and $\perm \in \{\rwlx,\rwl\}$, then $\gl = \local$.
    \end{itemize}
  \end{lemma}
  In our definition of worlds, nothing prevents a world from having regions that are overlapping.
  This may seem like an issue with the $\stdvr$-relation as it allows different permission-based requirements to be satisfied by different regions.
  In practice, it is not an issue as we will always have a memory satisfaction assumption which doubles as a well-formedness condition on the world as it prevents the regions from overlapping.
  The memory satisfaction assumption in Lemma~\ref{lem:wl-local} is there to ensure a well-formed world.
\end{toplas}

@@

  In addition to these syntactic requirements, we also specify standard behavioural expectations of malloc.
  Intuitively, we require that when malloc is invoked with a length argument, then it will return a capability for a fresh piece of memory of that size.
  It should be fresh in the sense that malloc has not already allocated any of that memory and will not do so in the future.
  Also, when invoked with a nonsensical length argument such as a negative integer or a capability, malloc should simply fail.
  However, formulating these requirements is harder than one might expect.
  The problem is that a realistic implementation of malloc needs to rely on internal state (the free memory is part of malloc's internal state) that changes after every invocation and relies on invariants on that state.
  We can only expect that malloc behaves according to its specification if its internal state satisfies the current state of the invariants in an executing system.
  We express this in terms of the semantic model from \sectionname~\ref{sec:logical-relation}.

@@

The macro \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} captures the
parts of the calling convention related to actually transferring control to
adversarial code. Specifically, it pushes the contents of the private registers,
$\overline{r_{\var{priv}}}$, to the stack and pushes the ``restoration'' code to
the stack. The restoration code is executed as the first thing upon return; it
restores the stack pointer and the old program counter. After the restoration
code is pushed to the stack, \texttt{scall
  $r$($\overline{r_{\var{args}}}$,$\overline{r_{\var{priv}}}$)} adjusts the pc
to point to the first instruction after the jump and pushes it to the stack. The
\texttt{scall}macro constructs a protected return pointer from a stack pointer by
adjusting it to point to the first instruction of the return code and
encapsulating it by restricting its permission to $\entry$. Next, the \texttt{scall}
macro reduces the range of authority of the stack pointer to the unused part of
the stack and clears it (as discussed in Section~\ref{sec:stack-and-return-pointer}). Finally, the \texttt{scall} macro clears
non-argument registers and jumps to $r$. Upon return after the restoration code
on the stack has been executed, the \texttt{scall} macro pops the restoration code from
the stack and restores the private state by popping it from the stack to the
appropriate registers. \figurename~\ref{fig:scall-impl} displays the
implementation of \texttt{scall} and the restoration code used by
\texttt{scall}. The implementation of \texttt{scall} uses some of the macros we
present next.

@@

\begin{toplas}
\section{Reasoning about programs on a capability machine}
\label{sec:reasoning}
% \itoplassug{
%   \begin{itemize}
%   \item A section with the lemmas we use for reasoning about programs
%     \begin{itemize}
%     \item lemma 58, scall works. Include description of how it handles the
%       common things in call and leaves the argumentation for callee code and
%       continuation.
%     \item lemma 59, malloc works. If this is included, we should probably also
%       include the malloc specification.
%     \item lemma 60, create closure works.
%     \item Include a description about how the FTLR is used to reason about
%       unknown code.
%     \item Mention that this is a step on the way towards having a program logic,
%       and then in the discussion discuss why a program logic for this language
%       is not what we need.
%     \end{itemize}
%   \end{itemize}
% }
There are many details to get right when programming in assembly.
These details carry over to proofs about assembly programs, so many of the proofs in \sectionname~\ref{sec:examples} about example programs are a bit cumbersome.
It is especially annoying when the same line of reasoning is applied in multiple places.
To mitigate this, we have defined a number of lemmas that capture common reasoning patterns in these proofs.
In this section, we present the most central lemmas used to reason about assembly programs.

@@

% Many details to get right when reasoning about low-level programs.
Another common pattern in proofs about programs on our capability machine concerns the use of \texttt{scall}.
The following lemma captures the commonalities of reasoning about programs using \texttt{scall}.
For instance, setting up the local stack frame and constructing stack pointers and protected return pointers for the callee always amount to the same line of reasoning which is captured by the lemma.
From another point of view, it can be seen as a specification for \texttt{scall}.
\begin{lemma}[$\mathtt{scall}$ works]
  \label{lem:scall-works}
  For all $n \in \nats$, $\ms, \ms_\stk, \ms_\var{unused}, \ms_f \in \MemSegments$, $W
  \in \Worlds$, $\reg \in \Regs$, $r,\overline{r_{\mathit{arg}}},
  \overline{r_{\mathit{priv}}} \in \RegName$, and $c_\var{next} \in \Caps$,
  if
  \begin{enumerate}
  \item \label{item:scall:memsat} $\memSat[n]{\ms}{\revokeTemp{W}}$ 
  \item $\dom(\ms_f) \cap (\dom(\ms_\stk \uplus \ms_{\mathit{unused}} \uplus \ms)) = \emptyset$
  \item \label{item:scall:stack} $(\reg,\ms) \text{ is looking at }
    \mathtt{scall}\;r(\overline{r_{\mathit{arg}}},
    \overline{r_{\mathit{priv}}}) \text{ followed by }
    c_{\mathit{next}}$
  \item $\reg \text{ points to stack with $\ms_\stk$ used and $\ms_{\mathit{unused}}$ unused}$
  \item \label{item:scall:hyp-callee} \emph{Hyp-Callee} For all  $\ms_\var{rec},
    \ms_\var{unused}', \ms'' \in \MemSegments$, $W' \in \Worlds$, $c_\var{ret},
    c_\stk \in \Caps$, and $\reg' \in \Regs$, if
    \begin{itemize}
    \item $\dom(\ms_{\mathit{unused}}) = \dom(\ms_{\mathit{rec}}
      \uplus \ms_{\mathit{unused}}')$,
    \item $W' =
      \revokeTemp{W}[\iota^{\sta}(\temp,\ms_\stk\uplus\ms_{\mathit{rec}} \uplus \ms_f),\iota^{\pwl}(\dom(\ms_{\mathit{unused}}'))]$\footnote{We use the update-notation without a region name to indicate that it should be a fresh region name.},
    \item $\memSat[n-1]{\ms''}{W'}$
    \item $\reg' \text{ points to stack with $\emptyset$ used and $\ms_{\mathit{unused}}'$ unused}$
    \item $\reg'= \reg_0[\pcreg\mapsto\updatePcPerm{\reg(r)},
      \overline{r_{\mathit{arg}}} \mapsto \reg(\overline{r_{\mathit{arg}}}),r_0
      \mapsto c_{\mathit{ret}}, r_\stk \mapsto c_\stk, r \mapsto \reg(r)]$ 
    \item $\npair[n-1]{c_{\mathit{ret}}} \in \stdvr(W')$
    \item $\npair[n-1]{c_\stk} \in \stdvr(W')$
    \end{itemize}

@@

In the common case, \texttt{scall} is used to reason about untrusted code that we only have general assumptions about.
As explained previously, we use the FTLR (Theorem~\ref{thm:ftlr}) to argue about unknown code; this is no exception.
That is, when we argue that assumption~\ref{item:scall:hyp-callee} in the \texttt{scall} lemma is satisfied, we use the FTLR along with the assumptions we have (e.g.\ linking table, malloc capability etc.).
In Hyp-Callee, we assume the memory is safe, so it suffices to show that the register-file contains safe values, which amounts to showing that the arguments in the call are safe.

@@

% We stress that these lemmas
% are conveniences and are not strictly necessary to reason about programs for the
% capability machine.
% In principle, one could reason solely in the semantic model
% without the use of these lemmas, but that would lead to a lot of repetition.
\end{toplas}

@@

The assertion on line 8 of Figure~\ref{fig:awkward} should never fail.
From the code, this seems natural since \texttt{adv} does not have access to $x$.
Therefore after the second invocation on line 7, $x$ should still be in the state it was before that invocation.
However, \texttt{adv} could have access to $\textit{cl}_x$ and could invoke $\textit{cl}_x$ again within the second invocation of \texttt{adv}.
Then $\textit{cl}_x$ would set $x$ to $0$ again (on line 5) and reinvoke another callback.
If that second callback were somehow able to skip its own caller and return directly to the caller of \texttt{adv}, it would end up on line 8 of Figure~\ref{fig:awkward} with $x$ set to $0$, causing the assertion to fail.
Without going into details, such an attack is perfectly possible if \texttt{adv} has access to a call/cc primitive (or equivalent).
In other words, if we are able to prove that the assertion will never fail, this attack, and other similar attacks, on well-bracketed control flow are adequately prevented.

@@

  The next part in the execution is an \texttt{scall}, so, according to \sectionname~\ref{sec:reasoning}, we should apply the \texttt{scall} lemma (Lemma~\ref{lem:scall-works}).
  For the sake of presentation, we here sketch some of the things the \texttt{scall} lemma actually takes care of.
  Based on the stack pointer's permission, we know by Lemma~\ref{lem:wl-local} that the capability is local (because it is a system wide assumption that there are no global read/write-local/execute capabilities \textit{cf.}\ \sectionname~\ref{sec:stack-and-return-pointer}) which means that the region that governs it must be temporary.
  This allows us to construct a world $W_2 \futurestr W_1$ where the stack region of $W_1$ has been revoked.
  In $W_2$, two new regions govern the stack.
  One of the two new regions governs the caller stack frame, i.e.\ the part of the stack that contains the contents of private registers and stack recovery code.
  The region that takes care of this is a temporary static region which ensures that our local stack frame remains unchanged during the execution of the callback.
  The other new region takes care of the unused part of the stack which we are going to let the callback use for its execution.
  This is taken care of by a standard region $\iota^{pwl}$-region which allows the callback to store any safe value on the stack.
  The callback is global, so it is safe to invoke it in a private future world of $W_1$ (even though the code is unknown, we do not need to use the FTLR because we assume that the arguments for the invocation of $c_{f4}$ are safe).
  Notice that had the callback been local, this would not be the case, but it would also not be safe to invoke as it might be a stack pointer as discussed in \sectionname~\ref{sec:stack-and-return-pointer}.
  Note also that arguing that the memory satisfies $W_2$ when we invoke the callback only works because we cleared the stack entirely (including the unused part) before the invocation.
  Otherwise, it might contain local values that are only known to be safe in $W_1$, but for which we do not know that they will remain safe in the private future world $W_2$.
  This part corresponds to arguing about Hyp-Callee in the \texttt{scall} lemma.

@@

\section{Discussion}
\label{sec:discussion}
\itoplassug{Reviewer C, esop, would have liked a discussion of (a) what could be done on a machine with a smaller set of capabilities, e.g., no local capability (we could probably not have done anything as the stack pointer could be stored on the heap and reused at unintended times.) and (b) what could be done with a larger/stronger set of available capabilities (we could include a discussion of linear capabilities).}
\itoplassug{Reviewer C, esop, asks whether our attacker model is reasonable (see tex comment). Maybe we should include a short discussion of the attacker model.}
% p.18 "they are linked with an adversary, adv, that is allowed to allocate memory but has no other capabilities".  Is this a reasonable model of the adversary?  (Genuinely naive question.)
\itoplassug{Should we include a discussion of how this ``scales'' to other things like a multi-core setting or if we needed tail calls?}
% * Also, the local capabilities seem related to Boyland's "Borrowing"
%   e.g. that is now supported in Rust amongst other research languages
%   (Pony etc) also mentioned in the ECOOP paper.
% * the combination of E and L capabilities seems very interesting
%   indeed.  I wonder if e.g. they could be used to support strong
%   models of encapsulation e.g. like ownership types. (see eg LNCS
%   7850)
\noindent\textbf{Calling convention}\\
\emph{Formulating control flow correctness} While we claim that our calling convention enforces control-flow correctness, we do not prove a general theorem that shows this because it is not clear what such a theorem should look like.
Formulations in terms of a control-flow graph, like the one by \citet{abadi_control-flow_2005}, \itoplas{creates a view with all capabilities that may be available at some point in time.
By removing the time aspect, we may have to consider capabilities that are actually not available at a given point in time.}
  Example~\texttt{g1} relies on such temporal properties when returning from the higher-order call backs.}
In fact, our examples show that our logical relation implies a stronger form of control-flow correctness than such formulations, although this is not made very explicit.
In later work, \citet{skorstengaard_stktokens_2019} provide a more explicit and useful definition of control-flow correctness and local state encapsulation.
The idea of their definition is to define a variant of a simple capability machine with a built-in stack as well as call and return instructions.
The call and return instructions use the built-in stack which gives well-bracketed control and local state encapsulation by construction.
They prove full abstraction~\citep{abadi_protection_1998} for a compilation from the capability machine with the built-in stack to one without which means that the programs running on the capability machine without a built-in stack behave as though there was a built-in stack.
Their full abstraction proof uses a logical relation, but it differs in a number of ways from the one we present.
Their logical relation is a binary cross language logical relation tailored specifically to prove the full abstraction result.
This means that it could not be used to prove program correctness results similar to ours.
Their calling convention is based on something called linear capabilities rather than local capabilities.
Linear capabilities offer a different kind of limited revocation.
\itoplassug{Maybe elaborate on why full-abstractness and not weaker properties like ``robust safety''.}
